---
title: "The effects of a severe wildfire on a bat community in the Canadian Rocky Mountains"
author: "Katie Florko"
date: "25/01/2024"
output:
  html_document:
    toc: yes
    toc_float: yes
    toc_depth: 3
    number_sections: yes
    theme: united
    highlight: tango
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Overview

This document contains the code for the analyses associated with "*The effects of a severe wildfire on a bat community in the Canadian Rocky Mountains*". 

# Introduction

## Libraries

```{r}
library(tidyr)
require(ggplot2) # this is for data viz
require(tidyverse) # this is for data wrangling
require(lubridate) # for formatting dates
require(hms) # for formatting time
require(car) # for plotting the results with Component+Residual and doing vif test
require(effects) # for plotting the modelled effect of a covariate 
require(MuMIn) # for model comparison
require(lme4) # to run glmer
require(gridExtra) #arranging plots
require(cowplot) #arranging plots
require(ggeffects) #for the glm prediction plots
require(fastDummies) #for creating dummy variables (characters to numeric)
require(ggrepel) # for labelling PCA plot
library(dplyr)
library(jtools)
library(glmmTMB)
library(tibble)
```

# Data wrangling
## Load data

```{r}
#set working directory **ERIN** add your own line here, or go Session > Set Working Directory > Choose Directory, choose where your data files are, and then copy the line in your console into here
setwd("~/Documents/R/elow_bats")

bat <- read.csv("SpeciesConsolidated_FINAL_updated5.csv", header = TRUE)
head(bat)

```

## Reformat data 
```{r}
bats <- bat %>%
  mutate(Date = ymd(Date), # set date to be in a R recognized date format
         BatSunset = as_hms(BatSunset), # set in a R recognized time format
         BatSunrise = as_hms(BatSunrise)) %>%
  tidyr::spread(key = SpeciesConsolidated, value = Count) %>% # puts each species as a column
  mutate(totalbats = rowSums(dplyr::select(.,m30k:mMYEV)), # creates a sum column for all bats
         HighF = m40kMyotis + mHighF + mLABO + mMYEV, #creates a sum for HighF
         LowF = mEPFULANO +mLACI + mLowF) #creates a sum for LowF
  
```


# GLMS

## - Overdis function
```{r}

#add function to test for over dispersion in the glms
overdisp_fun <- function(model) {
    rdf <- df.residual(model)
    rp <- residuals(model,type="pearson")
    Pearson.chisq <- sum(rp^2)
    prat <- Pearson.chisq/rdf
    pval <- pchisq(Pearson.chisq, df=rdf, lower.tail=FALSE)
    c(chisq=Pearson.chisq,ratio=prat,rdf=rdf,p=pval)
}

```

## - Prep data
```{r}
# add count of site visits
sitevis <- bats %>%
  group_by(SiteName)%>%
  summarise(visits.n = n())

# merge site visits and format
bats22 <- bats %>%
  mutate(Date = as.Date(Date, format = "%Y-%m-%d")) %>%
  as_tibble() %>%
  # create BACI columns
  mutate(BA = case_when(PrePost == "Pre" ~ "before",
                        PrePost == "Post" ~ "after"),
         CI = case_when(BurnUnburn.N == "Unburn" ~ "control",
                        BurnUnburn.N == "Burn" ~ "impact")) %>%
  left_join(sitevis, by = c("SiteName" = "SiteName")) %>%
  mutate(ForestType = as.factor(ForestType), #factors for categorical data
         HabitatType = as.factor(HabitatType),
         NearestWaterFeature = factor(NearestWaterFeature, 
                                      levels=c("Creek","Wetland","Lake", "River")),
         SiteName = as.factor(SiteName),
         totalbats = as.integer(totalbats)) %>%
  mutate(NightSumPrecipp = scale(NightSumPrecip),
         SunsetTempp = scale(SunsetTemp),
         DayOfYearr = scale(DayOfYear))

# set baselines
bats22$BA <- factor(bats22$BA, levels = c("before", "after"))
bats22$CI <- factor(bats22$CI, levels = c("control", "impact"))


# see counts
## Year 1 post-fire
bats22_2018 <- bats22 %>%
  filter(Year < 2019)
table(bats22_2018$Year)


## Year 2 post-fire
bats22_2019 <- bats22 %>%
  filter(Year != 2018 &
           Year != 2020)
table(bats22_2019$Year)

## Year 2 post-fire
bats22_2020 <- bats22 %>%
  filter(Year != 2018 &
           Year != 2019)
table(bats22_2020$Year)

```


## - Fit models
### -- m40kMyotis
```{r}
mod.40k <- glmer.nb(m40kMyotis ~ BA*CI + 
                          NightSumPrecipp + 
                          SunsetTempp +
                          DayOfYearr + # didn't turn out to be important
                          (1|SiteName),
                 data = bats22, # the data with scaled variables
                 na.action = "na.fail") 

# see model summary
summary(mod.40k) 

# see model R2
performance::r2(mod.40k) # site explains a lot of the variability


## checking diagnostics
# check residuals:
qqnorm(resid(mod.40k)) #
qqline(resid(mod.40k)) # looks good


#check singularity:
tt <- getME(mod.40k,"theta")
ll <- getME(mod.40k,"lower")
min(tt[ll==0]) # should be larger tha 0.05 - looks good

# check dispersion
overdisp_fun(mod.40k) #ratio should be between 0.1 and 5 to be okay

# check covariance
car::vif(mod.40k) # looking at GVIF (remove values >3); do one at a time

# do model selection
subset(MuMIn::dredge(mod.40k), delta < 4) # best-supported model includes BA*CI + NSP + SnT

# refit best model
mod.40k2  <- glmer.nb(m40kMyotis ~ BA*CI + 
                          NightSumPrecipp +
                          SunsetTempp + # didn't turn out to be important
                          (1|SiteName),
                 data = bats22, # the data with scaled variables
                 na.action = "na.fail") 

# summary
summary(mod.40k2)
performance::r2(mod.40k2) # site explains a lot of the variability


# plots
plot(ggeffects::ggpredict(mod.40k2, terms = c("NightSumPrecipp[all]")))
plot(ggeffects::ggpredict(mod.40k2, terms = c("SunsetTempp[all]")))
plot(ggeffects::ggpredict(mod.40k2, terms = c("BA", "CI")))


##----- Year 1
mod.40k2018  <- glmer.nb(m40kMyotis ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2018, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mod.40k2018), delta < 2) 

# refit best model
m40k2018_results <- summary(glmer.nb(m40kMyotis ~ BA*CI + NightSumPrecipp +
                 (1|SiteName),data = bats22_2018, na.action = "na.fail"))$coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2018", Species = "m40kMyotis")

plotdat <- ggeffects::ggpredict(glmer.nb(m40kMyotis ~ BA*CI + NightSumPrecipp +
                 (1|SiteName),data = bats22_2018, na.action = "na.fail"), 
                 terms = c("BA", "CI"))

plot_m40k2018_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted, color = group, 
                ymin = conf.low, ymax = conf.high), 
                position=position_dodge(0.2), width = 0.2, size = 0.75) +
  geom_point(data = plotdat, aes(x = x, y = predicted, color = group, shape =group),
             position=position_dodge(0.2), size = 2) + 
  geom_line(data = plotdat, aes(x = x, y = predicted, color = group, 
            group = group), position=position_dodge(0.2), size = 0.75) +
  scale_color_manual(values = c("forestgreen", "black")) +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 1 post-fire\n(2018)")) +
  scale_shape_manual(values = c(16, 4), breaks=c("control","impact"))  +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,160)


##----- Year 2
mod.40k2019  <- glmer.nb(m40kMyotis ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2019, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mod.40k2019), delta < 2) 

# refit best model
m40k2019_results  <- summary(glmer.nb(m40kMyotis ~ NightSumPrecipp + (1|SiteName),
                              data = bats22_2019, na.action = "na.fail")) $coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2019", Species = "m40kMyotis")

plot_m40k2019_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted, color = group, 
                ymin = conf.low, ymax = conf.high), 
                position=position_dodge(0.2), 
                width = 0.2, size = 0.75, alpha = 0) +
  geom_point(data = plotdat, aes(x = x, y = predicted, color = group),
             position=position_dodge(0.2), size = 2, alpha = 0) + 
  geom_line(data = plotdat, aes(x = x, y = predicted, color = group, 
            group = group), 
            position=position_dodge(0.2), size = 0.75, alpha = 0) +
  scale_color_manual(values = c("forestgreen", "black")) +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 2 post-fire\n(2019)")) +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,160)



##----- Year 3
mod.40k2020  <- glmer.nb(m40kMyotis ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2020, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mod.40k2020), delta < 2) 

# refit best model
m40k2020_results  <- summary(glmer.nb(m40kMyotis ~ BA*CI + DayOfYearr 
                            + NightSumPrecipp + (1|SiteName),
                            data = bats22_2020, na.action = "na.fail")) $coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2020", Species = "m40kMyotis")

plotdat <- ggeffects::ggpredict(glmer.nb(m40kMyotis ~ BA*CI + 
                                           DayOfYearr + NightSumPrecipp +
                 (1|SiteName),data = bats22_2020, na.action = "na.fail"), 
                 terms = c("BA", "CI"))

plot_m40k2020_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted, color = group, 
                ymin = conf.low, ymax = conf.high), 
                position=position_dodge(0.2), width = 0.2, size = 0.75) +
  geom_point(data = plotdat, aes(x = x, y = predicted, color = group, shape = group),
             position=position_dodge(0.2), size = 2) + 
  geom_line(data = plotdat, aes(x = x, y = predicted, color = group, 
            group = group), position=position_dodge(0.2), size = 0.75) +
  scale_color_manual(values = c("forestgreen", "black")) +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 3 post-fire\n(2020)")) +
  scale_shape_manual(values = c(16, 4), breaks=c("control","impact"))  +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,160)


```


### -- mEPFULANO
```{r}
mod.mEPFULANO <- glmer.nb(mEPFULANO ~ BA*CI + 
                          NightSumPrecipp + # didn't turn out to be important
                          SunsetTempp +
                          DayOfYearr + 
                          (1|SiteName),
                 data = bats22, # the data with scaled variables
                 na.action = "na.fail") 

# see model summary
summary(mod.mEPFULANO) 

# see model R2
performance::r2(mod.mEPFULANO) # site explains a lot of the variability


## checking diagnostics
# check residuals:
qqnorm(resid(mod.mEPFULANO)) #
qqline(resid(mod.mEPFULANO)) # looks good


#check singularity:
tt <- getME(mod.mEPFULANO,"theta")
ll <- getME(mod.mEPFULANO,"lower")
min(tt[ll==0]) # should be larger than 0.05 - looks good

# check dispersion
overdisp_fun(mod.mEPFULANO) #ratio should be between 0.1 and 5 to be okay

# check covariance
car::vif(mod.mEPFULANO) # looking at GVIF (remove values >3); do one at a time

# do model selection
subset(MuMIn::dredge(mod.mEPFULANO), delta < 4) # best-supported model includes BA*CI + DOY + SnT


# plot
plot(ggeffects::ggpredict(mod.mEPFULANO, terms = c("BA[all]", "CI")))


# refit best model
mod.mEPFULANO2  <- glmer.nb(mEPFULANO ~ BA*CI + 
                          DayOfYearr +
                          SunsetTempp + # didn't turn out to be important
                          (1|SiteName),
                 data = bats22, # the data with scaled variables
                 na.action = "na.fail") 

# summary
summary(mod.mEPFULANO2)
performance::r2(mod.mEPFULANO2)

# plots
plot(ggeffects::ggpredict(mod.mEPFULANO2, terms = c("DayOfYearr[all]")))
plot(ggeffects::ggpredict(mod.mEPFULANO2, terms = c("SunsetTempp[all]")))
plot(ggeffects::ggpredict(mod.mEPFULANO2, terms = c("BA", "CI")))


##----- Year 1
mEPFULANO2018  <- glmer.nb(mEPFULANO ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2018, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mEPFULANO2018), delta < 2) 

# refit best model
mEPFULANO2018_results <- summary(glmer.nb(mEPFULANO ~ 
                 BA*CI +DayOfYearr + SunsetTempp +
                 (1|SiteName), data = bats22_2018, na.action = "na.fail"))$coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2018", Species = "mEPFULANO")

plotdat <- ggeffects::ggpredict(glmer.nb(mEPFULANO ~ BA*CI + 
                                           DayOfYearr + SunsetTempp +
                 (1|SiteName),data = bats22_2018, na.action = "na.fail"), 
                 terms = c("BA", "CI"))

plot_mEPFULANO2018_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted, color = group, 
                ymin = conf.low, ymax = conf.high), 
                position=position_dodge(0.2), width = 0.2, size = 0.75) +
  geom_point(data = plotdat, aes(x = x, y = predicted, color = group, shape = group),
             position=position_dodge(0.2), size = 2) + 
  geom_line(data = plotdat, aes(x = x, y = predicted, color = group, 
            group = group), position=position_dodge(0.2), size = 0.75) +
  scale_color_manual(values = c("forestgreen", "black")) +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 1 post-fire\n(2018)")) +
  scale_shape_manual(values = c(16, 4), breaks=c("control","impact"))  +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,50)




##----- Year 2
mEPFULANO2019  <- glmer.nb(mEPFULANO ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2019, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mEPFULANO2019), delta < 2) 

# refit best model
mEPFULANO2019_results  <- summary(glmer.nb(mEPFULANO ~ 
                              BA + DayOfYearr + SunsetTempp + (1|SiteName),
                              data = bats22_2019, na.action = "na.fail")) $coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2019", Species = "mEPFULANO")


plotdat <- ggeffects::ggpredict(glmer.nb(mEPFULANO ~ BA + 
                                           DayOfYearr + SunsetTempp +
                 (1|SiteName),data = bats22_2019, na.action = "na.fail"), 
                 terms = c("BA"))

plot_mEPFULANO2019_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted,  
                ymin = conf.low, ymax = conf.high), 
                width = 0.2, size = 0.75, color = "grey70") +
  geom_point(data = plotdat, aes(x = x, y = predicted), 
             size = 2, color = "grey70") + 
  geom_line(data = plotdat, aes(x = x, y = predicted, 
            group = group), size = 0.75, color = "grey70") +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 2 post-fire\n(2019)")) +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,50)






##----- Year 3
mEPFULANO2020  <- glmer.nb(mEPFULANO ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2020, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mEPFULANO2020), delta < 2) 

# refit best model
mEPFULANO2020_results  <- summary(glmer.nb(mEPFULANO ~ BA + DayOfYearr 
                            + SunsetTempp + (1|SiteName),
                            data = bats22_2020, na.action = "na.fail")) $coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2020", Species = "mEPFULANO")



plotdat <- ggeffects::ggpredict(glmer.nb(mEPFULANO ~ BA + 
                                           DayOfYearr + SunsetTempp +
                 (1|SiteName),data = bats22_2020, na.action = "na.fail"), 
                 terms = c("BA"))

plot_mEPFULANO2020_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted,  
                ymin = conf.low, ymax = conf.high), 
                width = 0.2, size = 0.75, color = "grey70") +
  geom_point(data = plotdat, aes(x = x, y = predicted), 
             size = 2, color = "grey70") + 
  geom_line(data = plotdat, aes(x = x, y = predicted, 
            group = group), size = 0.75, color = "grey70") +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 3 post-fire\n(2020)")) +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,50)



```


### -- mLACI
```{r}
mod.mLACI <- glmer.nb(mLACI ~ BA*CI + 
                          NightSumPrecipp + 
                          SunsetTempp +
                          DayOfYearr + # didn't turn out to be important
                          (1|SiteName),
                 data = bats22, # the data with scaled variables
                 na.action = "na.fail") 

# see model summary
summary(mod.mLACI) 

# see model R2
performance::r2(mod.mLACI) # site explains a lot of the variability




## checking diagnostics
# check residuals:
qqnorm(resid(mod.mLACI)) #
qqline(resid(mod.mLACI)) # looks good


#check singularity:
tt <- getME(mod.mLACI,"theta")
ll <- getME(mod.mLACI,"lower")
min(tt[ll==0]) # should be larger than 0.05 - looks good

# check dispersion
overdisp_fun(mod.mLACI) #ratio should be between 0.1 and 5 to be okay

# check covariance
car::vif(mod.mLACI) # looking at GVIF (remove values >3); do one at a time

# do model selection
subset(MuMIn::dredge(mod.mLACI), delta < 2) # best-supported model includes BA*CI + DOY + SnT

# refit best model
mod.mLACI2 <- glmer.nb(mLACI ~ BA + 
                          SunsetTempp +
                          DayOfYearr + # didn't turn out to be important
                          (1|SiteName),
                 data = bats22, # the data with scaled variables
                 na.action = "na.fail") 

# summary
summary(mod.mLACI2)
performance::r2(mod.mLACI2)


# plots
plot(ggeffects::ggpredict(mod.mLACI2, terms = c("SunsetTempp[all]")))
plot(ggeffects::ggpredict(mod.mLACI2, terms = c("DayOfYearr[all]")))
plot(ggeffects::ggpredict(mod.mLACI2, terms = c("BA")))


##----- Year 1
mLACI2018  <- glmer.nb(mLACI ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2018, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mLACI2018), delta < 2) 

# refit best model
mLACI2018_results <- summary(glmer.nb(mLACI ~ 
                 BA*CI +DayOfYearr + SunsetTempp +
                 (1|SiteName),data = bats22_2018, na.action = "na.fail"))$coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2018", Species = "mLACI")


plotdat <- ggeffects::ggpredict(glmer.nb(mLACI ~ 
                 BA*CI +DayOfYearr + SunsetTempp +
                 (1|SiteName), data = bats22_2018, na.action = "na.fail"), 
                 terms = c("BA", "CI"))

plot_mLACI2018_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted, color = group, 
                ymin = conf.low, ymax = conf.high), 
                position=position_dodge(0.2), width = 0.2, size = 0.75) +
  geom_point(data = plotdat, aes(x = x, y = predicted, color = group, shape = group),
             position=position_dodge(0.2), size = 2) + 
  geom_line(data = plotdat, aes(x = x, y = predicted, color = group, 
            group = group), position=position_dodge(0.2), size = 0.75) +
  scale_color_manual(values = c("forestgreen", "black")) +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 1 post-fire\n(2018)")) +
  scale_shape_manual(values = c(16, 4), breaks=c("control","impact"))  +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,20)




##----- Year 2
mLACI2019  <- glmer.nb(mLACI ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2019, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mLACI2019), delta < 2) 

# refit best model
mLACI2019_results  <- summary(glmer.nb(mLACI ~ 
                              BA + DayOfYearr + SunsetTempp + (1|SiteName),
                              data = bats22_2019, na.action = "na.fail")) $coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2019", Species = "mLACI")


plotdat <- ggeffects::ggpredict(glmer.nb(mLACI ~ BA + 
                                           DayOfYearr + SunsetTempp +
                 (1|SiteName),data = bats22_2019, na.action = "na.fail"), 
                 terms = c("BA"))

plot_mLACI2019_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted,  
                ymin = conf.low, ymax = conf.high), 
                width = 0.2, size = 0.75, color = "grey70") +
  geom_point(data = plotdat, aes(x = x, y = predicted), 
             size = 2, color = "grey70") + 
  geom_line(data = plotdat, aes(x = x, y = predicted, 
            group = group), size = 0.75, color = "grey70") +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 2 post-fire\n(2019)")) +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,20)





##----- Year 3
mLACI2020  <- glmer.nb(mLACI ~ BA*CI + NightSumPrecipp +SunsetTempp + 
                         DayOfYearr + (1|SiteName),data = bats22_2020, 
                         na.action = "na.fail") 
subset(MuMIn::dredge(mLACI2020), delta < 2) 

# refit best model
mLACI2020_results  <- summary(glmer.nb(mLACI ~ BA + DayOfYearr + 
                            SunsetTempp + (1|SiteName),
                            data = bats22_2020, na.action = "na.fail")) $coef %>%
  as.data.frame() %>%
  rownames_to_column("Terms") %>%
  as_tibble() %>%
  mutate(Year = "2020", Species = "mLACI")

plotdat <- ggeffects::ggpredict(glmer.nb(mLACI ~ BA + 
                                           DayOfYearr + SunsetTempp +
                 (1|SiteName),data = bats22_2020, na.action = "na.fail"), 
                 terms = c("BA"))

plot_mLACI2020_results <- ggplot() +
  geom_errorbar(data = plotdat, aes(x = x, y = predicted,  
                ymin = conf.low, ymax = conf.high), 
                width = 0.2, size = 0.75, color = "grey70") +
  geom_point(data = plotdat, aes(x = x, y = predicted), 
             size = 2, color = "grey70") + 
  geom_line(data = plotdat, aes(x = x, y = predicted, 
            group = group), size = 0.75, color = "grey70") +
  scale_x_discrete(labels= c("Years before fire\n(2015-2017)", "Year 3 post-fire\n(2020)")) +
  theme_minimal() + theme(legend.position = "none",
                          text = element_text(size=12)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  ggtitle(NULL) + ylab(" ") + xlab("") + ylim(0,20)


```


# Plots
## - 1. Plot raw data
```{r}

bats2 <- bats %>% 
  as_tibble() %>%
  # create BACI columns
  mutate(BA = case_when(PrePost == "Pre" ~ "before",
                        PrePost == "Post" ~ "after"),
         CI = case_when(BurnUnburn.N == "Unburn" ~ "control",
                        BurnUnburn.N == "Burn" ~ "impact")) %>%
  pivot_longer(cols = c("m40kMyotis", "mEPFULANO", "mLACI"), names_to = "species", values_to = "count") %>%
  group_by(BA, CI, species) %>%
  summarise(meancount = mean(count),
            sd = sd(count),
            se = sd/sqrt(n()))

bats2$BA <- factor(bats2$BA, levels=c("before", "after"))


## plot!
plot_baci <- ggplot(bats2) +
  geom_point(aes(x = BA, y = meancount, color = CI, shape =CI), position=position_dodge(0.1)) +
  geom_line(aes(x = BA, y = meancount, group = CI, color = CI), position=position_dodge(0.1)) +
  geom_errorbar(aes(x = BA, ymin=meancount-se, ymax=meancount+se, group = CI, color = CI), width=.2, position=position_dodge(0.1)) +
  facet_wrap(~species, labeller = as_labeller(c(`m40kMyotis` = "A. 40k Myotis",
             
                    `mEPFULANO` = "B. Big brown & silver-haired bats",
                    `mLACI` = "C. Hoary bats")), ncol=3, scales = "free") +
  scale_colour_manual(values = c("forestgreen", "black"), breaks=c("control","impact"), labels = c("Not burned", "Burned"), name = "Sites that after the\n2017 Kenow Wildfire were:") +
  scale_shape_manual(values = c(16, 4), breaks=c("control","impact"), labels = c("Not burned", "Burned"), name = "Site status after the\n2017 Kenow Wildfire:") +
  ylab("Mean number of passes per night") +
  scale_x_discrete(labels= c("Before", "After")) +
  xlab("Before and after the wildlfire") +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10, hjust=1), 
        axis.text.x = element_text(size = 10),
        legend.background = element_rect(fill="white", colour = "grey90"),
        legend.text=element_text(size=10)) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1)

tiff("bats_baci.tiff", units="in", width=10, height=3.5, res=500)
plot_baci
dev.off()  
```


## - 2. by year
```{r}
#-------- prep data
bats_plots <- bats22
bats_plots$SiteName<- factor(bats_plots$SiteName, levels = c("Dipper at Rowe","Red Rock","Yarrow Cabin", "Wishbone","Blakiston Roadside","Boundary Bay", "Sofa Burn", "Sewage Lagoon","Bison Paddock", "Belly Bend", "Cameron Lake", "Lone Lake","Lost Lake"))
bats_plots$BurnUnburn.N <- factor(bats_plots$BurnUnburn.N, levels = c("Burn", "Unburn"))


bats_plots <- bats_plots %>%
  mutate(Date = as.POSIXct(Date, format = "%Y-%m-%d"),
    Year = year(Date)) %>%
  pivot_longer(cols = c("m40kMyotis", "mEPFULANO", "mLACI"), names_to = "species", values_to = "count") %>%
  group_by(species, Year, BurnUnburn.N) %>%
  summarise(meancount = mean(count),
            sd = sd(count),
            se = sd/sqrt(n()))

#-------- plot!
plot_year <- ggplot(bats_plots, aes(Year, meancount)) +
  #geom_rect(aes(xmin=2017.2,xmax=2020.5,ymin=-Inf,ymax=Inf), 
            #fill = "grey90", alpha = 1, color = "grey90") +
  #geom_vline(xintercept=2017.2, color = "red", size = 3) +
  geom_point(aes(color = BurnUnburn.N, shape = BurnUnburn.N), size = 2) +
  geom_errorbar(aes(x = Year, ymin=meancount-se, ymax=meancount+se, group = BurnUnburn.N, color = BurnUnburn.N), width=.15) +
  geom_line(aes(color = BurnUnburn.N, group = BurnUnburn.N)) +
  facet_wrap(species~., labeller = as_labeller(c(`m40kMyotis` = "A. 40k Myotis",
                    `mEPFULANO` = "\nB. Big brown and silver-haired bats\n",
                    `mLACI` = "\nC. Hoary bats\n")), scales = "free", ncol=1) +
  scale_colour_manual(values = c("forestgreen", "black"), breaks=c("Unburn","Burn"), labels = c("Not burned", "Burned"), name = "Sites that after the\n2017 Kenow Wildfire were:") +
  scale_shape_manual(values = c(16, 4), breaks=c("Unburn","Burn"), labels = c("Not burned", "Burned"), name = "Sites that after the\n2017 Kenow Wildfire were:") +
  theme_minimal() +
  guides(colour=guide_legend(title.position="top", 
                                     title.hjust =0.5)) +
  theme(strip.text = element_text(size = 12, hjust = 0),
        legend.position = "bottom",
        legend.background = element_rect(fill="white", colour = "grey90"),
        legend.text=element_text(size=10),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.title.align=0.5,
        legend.box.just = "center") +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1) +
  labs(x="Year", y = "Mean number of bats detected per night") 

ggsave(
  plot = plot_year,
  filename = "bats_year.png",
  bg = "transparent",
  units="in", width=4, height=9
)

```

## - 3. coefficient plots
```{r}
level_order <- c('DayOfYearr', 'SunsetTempp', 'NightSumPrecipp', 'CIimpact', 'BAafter', "BAafter:CIimpact") 


bats_coef <- rbind(m40k2018_results, m40k2019_results, m40k2020_results, 
      mEPFULANO2018_results, mEPFULANO2019_results, mEPFULANO2020_results,
      mLACI2018_results, mLACI2019_results, mLACI2020_results) %>%
  rename("SE" = "Std. Error", 
         "p" = "Pr(>|z|)") %>%
  dplyr::select(-"z value") %>%
  filter(Terms != "(Intercept)") %>%
  mutate(Significant = case_when(p > 0.05 ~ "no",
                                 p < 0.05 ~ "yes")) %>%
  ggplot() +
  geom_point(aes(x = Estimate, y = factor(Terms, level = level_order), 
                 color = Year, shape = Year, alpha = Significant), position=position_dodge2(reverse = TRUE, 
                                                         width = 0.5), size = 2) +
  geom_errorbar(aes(xmin = Estimate-SE, xmax = Estimate+SE, y = Terms, 
                    color = Year, alpha = Significant), width = 0.5, 
                position=position_dodge2(reverse = TRUE, width = 0.9), size =0.75) +
  scale_color_manual(values = c("#160A37", "#BF4F4A", "#F1B543"), name = "Year (years\nsince fire)", breaks = c(2018, 2019, 2020), labels = c("2018 (1)", "2019 (2)", "2020, (3)")) +
  scale_shape_manual(values = c(15, 16, 17), name = "Year (years\nsince fire)", breaks = c(2018, 2019, 2020), labels = c("2018 (1)", "2019 (2)", "2020, (3)")) +
  scale_y_discrete(labels=c("Ordinal date", "Temperature", "Precipitation", "Burned or not burned", "Before or after wildfire", "Burned/not : before/after")) +
  scale_alpha_discrete(range = c(0.35, 0.9)) +
  ylab("Coefficient") +
  facet_grid(Year~Species, labeller = as_labeller(c(`m40kMyotis` = "A. m40k Myotis",
                    `mEPFULANO` = "\nB. Big brown and silver-haired bats\n",
                    `mLACI` = "\nC. Hoary bats\n",
                    '2018' = "2018",
                    '2019' = "2019",
                    '2020' = "2020"))) +
  theme_minimal() +
  theme(strip.text = element_text(size = 12, hjust = 0),
        legend.position = "right",
        legend.background = element_rect(fill="white", colour = "grey90"),
        legend.text=element_text(size=10),
        legend.title.align=0.5,
        legend.box.just = "center") +
  geom_vline(xintercept = 0, linetype = 2, size = .25) +
  annotate("segment", x=-Inf, xend=Inf, y=-Inf, yend=-Inf, size = 1) +
  annotate("segment", x=-Inf, xend=-Inf, y=-Inf, yend=Inf, size = 1)


tiff("bats_coef.tiff", units="in", width=11, height=6, res=500)
bats_coef
dev.off()  


```

## - 4. effect plots

```{r}
tiff("bats_effects.tiff", units="in", width=9, height=6, res=500)
plot_grid(plot_m40k2018_results, plot_mEPFULANO2018_results, plot_mLACI2018_results, plot_m40k2019_results, plot_mEPFULANO2019_results, plot_mLACI2019_results,
plot_m40k2020_results, plot_mEPFULANO2020_results, plot_mLACI2020_results, ncol = 3)
dev.off()
```

##- 5. description of sites
```{r}

bats2 <- bats %>%
  group_by(SiteName)%>%
  dplyr::summarize(meanbats = mean(totalbats),
            NearestWaterFeature=NearestWaterFeature,
            TownsiteDistance=TownsiteDistance,
            Elevation=Elevation,
            BurnUnburn.N=BurnUnburn.N)%>%
  distinct()

fig_S1 <- ggplot(bats2, aes(TownsiteDistance, Elevation)) +
  xlab("Townsite distance (km)") + ylab("Elevation (m)") +
  geom_point(aes(shape = NearestWaterFeature, size = meanbats, color = BurnUnburn.N), alpha = 0.9) + 
  labs(shape = "Nearest water feature", size = "Approximate mean\npasses / night", colour = "Burn or unburn") + 
  scale_size(breaks = c(50, 100, 250, 500, 750), labels = c("50", "100", "250", "500", "750")) + 
  theme_ggeffects() +
  geom_text_repel(aes(label = SiteName)) +
  guides(color = guide_legend(override.aes = list(size = 4)),
         size = guide_legend(override.aes = list(size = 2)),
         shape = guide_legend(override.aes = list(size = 4))) 


tiff("fig_s1.tiff", units="in", width=8, height=5, res=300)
fig_S1
dev.off()

```

# Session info
```{r}
sessioninfo::session_info()
```

